### 2019-06-07 paper submission 3tex.tex
Make easier edits in text.

The consumption bundle $C_{nt}$ is packaged by a consumption-good producer
using the Cobb-Douglas aggregate 
\begin{equation}
C_{nt}=\prod\nolimits_{j=1}^{J}\left( C_{nt}^{j}\right) ^{\alpha _{t}^{j}},
\label{aggregate}
\end{equation}%
where $C_{nt}^{j}$ is the quantity of sectoral good $j$ used for
consumption, and $\sum_{j=1}^{J}\alpha _{t}^{j}=1$. The $\alpha $s are
allowed to change over time to capture possible changes in tastes. \textbf{%
In Section \ref{Sces} we investigate the robustness of our results to a
specification of preferences in which the elasticity of substitution among
sectoral goods is not unitary.}

@@

Using (\ref{BC}) in the utility function we can solve for the sectoral labor
allocation:%
\begin{equation}
L_{nt}^{j}=\arg \max E_{t-1}\left[ \log \left( \frac{%
\sum_{j=1}^{J}w_{nt}^{j}L_{nt}^{j}}{P_{nt}}\right) \right]
,s.t.:\sum\nolimits_{j=1}^{J}L_{nt}^{j}=L_{nt},  \label{eq:log:utility}
\end{equation}%
\textbf{where }$E_{t-1}$\textbf{\ indicates the rational expectation over
the possible realizations of period }$t$\textbf{\ shocks. In particular, the
representative agent knows }$L_{nt}^{j}$\textbf{\ and }$\kappa _{nt}^{j}$%
\textbf{, as well as the previous values of the shock processes }$%
A_{nt-1}^{j}$\textbf{. The agent also knows the distribution of }$A_{nt}^{j}$%
\textbf{\ conditional on }$A_{nt-1}^{j}$\textbf{\ (which we estimate later
from data), and is therefore able to compute the rational expectation in (%
\ref{eq:log:utility})}.\footnote{%
Implicit in our formulation is the assumption that there is perfect
risk-sharing within a country, but no risk-sharing across countries. To
motivate the lack of risk-sharing across countries, see our earlier
discussion on the high comovement between consumption and output as well as
the high correlation between consumption and output volatility.}

@@

\textbf{The second stage of the solution uses (\ref{eq:FOC}) to find the
ex-ante shares }$L_{nt}^{j}/L_{nt}$\textbf{. Our solution method computes
the rational expectation in (\ref{eq:FOC}) by bootstrapping from the
empirical distribution of }$A_{nt}^{j}$\textbf{. In particular, we begin
with a choice of candidate values for the }$L_{nt}^{j}$\textbf{s, and draw a
large number of realizations of the }$A_{nt}^{j}$\textbf{s from their
empirical distributions (conditional on the }$A_{nt-1}^{j}$\textbf{s). For
each of these realizations, we compute the solution for the }$w_{nt}^{j}$%
\textbf{s from the system (\ref{shares})-(\ref{RC}), and then the term in
brackets on the right side of (\ref{eq:FOC}). The rational expectation is
then the average of the terms in brackets across all the bootstrapped
realizations. If this is (close enough to being) equal to the starting guess
for }$L_{nt}^{j}/L_{nt}$\textbf{\ we stop. Otherwise, we move to a new guess
for }$L_{nt}^{j}/L_{nt}$\textbf{. A more detailed explanation is provided in
the Appendix.}

The key theoretical outcome we are interested in is aggregate income
volatility, which we measure as the variance (or standard deviation, where
indicated), of real income deviations from country-specific trends. In turn,
real income in the model is given by total value added deflated by the
optimal expenditure-based price index, or $Y_{nt}=\frac{w_{nt}L_{nt}}{P_{nt}}%
.$ As discussed in the Introduction, these welfare-relevant measures of
income are expected to show first-order responses to changes in the terms of
trade, and hence in foreign productivities, endowments, or trade costs.%
\footnote{%
In contrast, if we were to deflate nominal GDP by using the CES price
aggregates of the sector-level variety baskets, we would retrieve the
Kehoe-Ruhl invariance of GDP\ to shocks to the terms of trade. It is
doubtful, however, that GDP as constructed by statistical agencies maps well
into this theoretical construct. They may measure the price of a
representative variety within each sector, the average price of an aggregate
variety basket, or a random sample of continuously used varieties. This
choice might also depend on the source country, as import price indexes are
computed differently from producer price indexes (Nakamura and Steinsson,
2008 QJE vs Nakamura and Steinsson, 2012 AER). In contrast, the CPI is
easier to map to our model, because consumers only consume J different final
goods, not a continuum of varieties. Our welfare-relevant price index, which
is the geometric average of final good prices is a very close approximation
of the expenditure-weighted T\"{o}rnqvist price index, the way the CPI is
usually calculated.} \textbf{Please check this footnote very carefully.}

@@

\textbf{In order to solve the model, we must be able to compute the rational
expectation of a sector's value added as a share of total value added at the
beginning of period }$t$\textbf{, conditional on }$\hat{Z}_{nt-1}^{j}$%
\textbf{\ (see equation (\ref{eq:FOC})). To do this we assume that the
processes }$\lambda _{t}^{j}$\textbf{, }$\mu _{nt}$\textbf{, and }$\epsilon
_{nt}^{j}$\textbf{\ are all AR(1) log-normal processes, and for each of them
we estimate specific autoregressive and variance parameters. We then
construct possible realizations of }$Z_{nt}^{j}$\textbf{\ by bootsrapping
from the distributions of }$\lambda _{t}^{j}$\textbf{, }$\mu _{nt}$\textbf{,
and }$\epsilon _{nt}^{j}$\textbf{\ \ conditional on }$\lambda _{t-1}^{j}$%
\textbf{, }$\mu _{nt-1}$\textbf{, and }$\epsilon _{nt-1}^{j}$\textbf{, and
adding the bootstrapped values to the deterministic (trend) component of }$%
Z_{nt}^{j}$\textbf{.}

@@

Using (\ref{estimate}) in (\ref{AC}) we can solve for the ex-ante
allocation, which turns out to be independent of $\varrho $, and hence it
will be the same as in our baseline case, where $\varrho $ is implicitly
assumed to be infinite. In other words, $\psi _{nt_{-}}^{j}$ is given by
equation (\ref{eq:FOC}). \textbf{Miklos is this still true? I have vague
memories of conversations where we concluded this is not quite right. Have
we made any changes?}

@@

\subsubsection{CES Preferences}

\textbf{In our baseline model preferences over sectoral goods aggregate in
Cobb-Douglas fashion. In this robustness check we replace equation (\ref%
{aggregate}) by a CES formulation. This requires calibrating }$J$\textbf{\
\textquotedblleft share\textquotedblright\ parameters, as well as an
elasticity of substitution. Our strategy is to calibrate the share
parameters by .... and then look at how our results vary with different
values of the elasticity of substitution. }

\textbf{The results are presented in Table 6. The overall effect of trade on
volatility is quite similar across different specifications of preferences,
though when substitutability is quite low (elasticity of substitution equal
to 0.5) the diversification effect and the sectoral specialization effect
change in offsetting directions}

@@

@@

\textbf{The goal of this loop is to find the sectoral resource allocations }$%
L_{nt}^{j}$\textbf{\ that satisfy }%
\begin{equation*}
\frac{L_{nt}^{j}}{L_{nt}}=E_{t-1}\left( \frac{w_{nt}^{j}L_{nt}^{j}}{%
w_{nt}L_{nt}}\right)
\end{equation*}%
\textbf{where }$w_{nt}$\textbf{\ is the average wage. When searching for the
equilibrium value of }$L_{nt}^{j}$\textbf{\ the state of the economy is made
up of the deterministic component of the augmented-productivity processes, }$%
\bar{Z}_{nt}^{j}$\textbf{, as well as the previous-period values of the
log-deviation processes representing country, sector, and idyosincratic
shocks, }$\lambda _{t-1}^{j}$\textbf{, }$\mu _{nt-1}$\textbf{, and }$%
\epsilon _{nt-1}^{j}$\textbf{. This state is known both to us and to the
decision make in the model, as are the autoregressive parameters driving the
shock processes and their variances. Hence, we can draw values from the
distribution of }$\lambda _{t}^{j}$\textbf{, }$\mu _{nt}$\textbf{, and }$%
\epsilon _{nt}^{j}$\textbf{\ and combine them }$\bar{Z}_{nt}^{j}$\textbf{\
to create corresponding draws for }$Z_{nt}^{j}$\textbf{. For each iteration
over possible candidates for the }$L_{nt}^{j}$\textbf{s, we thus draw XXX
possible realizations of the }$Z_{nt}^{j}$\textbf{s, and for each of them we
compute }$w_{nt}^{j}$\textbf{\ and }$w_{nt}$\textbf{, and hence }$\frac{%
w_{nt}^{j}L_{nt}^{j}}{w_{nt}L_{nt}}$\textbf{\ from the middle loop. Then the
expectation of these wage shares is simply the average across all the draws
of }$Z_{nt}^{j}$\textbf{. The iteration ends when the left hand siode and
right hand side are close enough.}

@@

\textbf{Miklos: I have a feeling this needs to be updated}

>>>>

The consumption bundle $C_{nt}$ is packaged by a consumption-good producer
using the Cobb-Douglas aggregate 
\begin{equation}
C_{nt}=\prod\nolimits_{j=1}^{J}\left( C_{nt}^{j}\right) ^{\alpha _{t}^{j}},
\label{aggregate}
\end{equation}%
where $C_{nt}^{j}$ is the quantity of sectoral good $j$ used for
consumption, and $\sum_{j=1}^{J}\alpha _{t}^{j}=1$. The $\alpha $s are
allowed to change over time to capture possible changes in tastes. 
In Section \ref{Sces} we investigate the robustness of our results to a
specification of preferences in which the elasticity of substitution among
sectoral goods is not unitary.

@@

Using (\ref{BC}) in the utility function we can solve for the sectoral labor
allocation:%
\begin{equation}
L_{nt}^{j}=\arg \max E_{t-1}\left[ \log \left( \frac{%
\sum_{j=1}^{J}w_{nt}^{j}L_{nt}^{j}}{P_{nt}}\right) \right]
,s.t.:\sum\nolimits_{j=1}^{J}L_{nt}^{j}=L_{nt},  \label{eq:log:utility}
\end{equation}%
where $E_{t-1}$ indicates the rational expectation over
the possible realizations of period $t$ shocks. In particular, the
representative agent knows the previous values of the shock processes $%
A_{nt-1}^{j}$. The agent also knows the distribution of $A_{nt}^{j}$%
 conditional on $A_{nt-1}^{j}$ (which we estimate later
from data), and is therefore able to compute the rational expectation in (%
\ref{eq:log:utility}).\footnote{%
Implicit in our formulation is the assumption that there is perfect
risk-sharing within a country, but no risk-sharing across countries. To
motivate the lack of risk-sharing across countries, see our earlier
discussion on the high comovement between consumption and output as well as
the high correlation between consumption and output volatility.}

@@

The second stage of the solution uses (\ref{eq:FOC}) to find the
ex-ante shares $L_{nt}^{j}/L_{nt}$. Our solution method computes
the rational expectation in (\ref{eq:FOC}) by drawing from the
estimated distribution of $A_{nt}^{j}$. In particular, we begin
with a choice of candidate values for the $L_{nt}^{j}$s, and draw a
large number of realizations of the $A_{nt}^{j}$s from their
estimated distributions (conditional on the $A_{nt-1}^{j}$s). For
each of these realizations, we compute the solution for the $w_{nt}^{j}$%
s from the system (\ref{shares})-(\ref{RC}), and then the term in
brackets on the right side of (\ref{eq:FOC}). The rational expectation is
then the average of the terms in brackets across all the simulated
realizations. If this is (close enough to being) equal to the starting guess
for $L_{nt}^{j}/L_{nt}$ we stop. Otherwise, we move to a new guess
for $L_{nt}^{j}/L_{nt}$. A more detailed explanation is provided in
the Appendix.

The key theoretical outcome we are interested in is aggregate income
volatility, which we measure as the variance (or standard deviation, where
indicated), of real income deviations from country-specific trends. In turn,
real income in the model is given by total value added deflated by the
optimal expenditure-based price index, or $Y_{nt}=\frac{w_{nt}L_{nt}}{P_{nt}}%
.$ As discussed in the Introduction, these welfare-relevant measures of
income are expected to show first-order responses to changes in the terms of
trade, and hence in foreign productivities, endowments, or trade costs.%
\footnote{%
In contrast, if we were to deflate nominal GDP by using the CES price
aggregates of the sector-level variety baskets, we would retrieve the
Kehoe-Ruhl invariance of GDP to shocks to the terms of trade. It is
doubtful, however, that GDP as constructed by statistical agencies maps well
into this theoretical construct. They may measure the price of a
representative variety within each sector, the average price of an aggregate
variety basket, or a random sample of continuously used varieties. This
choice might also depend on the source country, as import price indexes are
computed differently from producer price indexes (Nakamura and Steinsson,
2008 and Nakamura and Steinsson, 2012). In contrast, the CPI is
easier to map to our model, because consumers only consume $J$ different final
goods, not a continuum of varieties. Our welfare-relevant price index, which
is the geometric average of final good prices is a very close approximation
of the expenditure-weighted T\"{o}rnqvist price index, the way the CPI is
usually calculated.}

@@

In order to solve the model, we must be able to compute the rational
expectation of a sector's value added as a share of total value added at the
beginning of period $t$, conditional on $\hat{Z}_{nt-1}^{j}$%
 (see equation (\ref{eq:FOC})). To do this we assume that the
processes $\lambda _{t}^{j}$, $\mu _{nt}$, and $\epsilon
_{nt}^{j}$ are all AR(1) log-normal processes, and for each of them
we estimate specific autoregressive and variance parameters. We then
construct possible realizations of $Z_{nt}^{j}$ by drawing
from the estimated distributions of $\lambda _{t}^{j}$, $\mu _{nt}$,
and $\epsilon _{nt}^{j}$ conditional on $\lambda _{t-1}^{j}$%
, $\mu _{nt-1}$, and $\epsilon _{nt-1}^{j}$, and
adding the drawn values to the deterministic (trend) component of log $%
Z_{nt}^{j}$.

@@

Using (\ref{estimate}) in (\ref{AC}) we can solve for the ex-ante
allocation. The first-order condition for $\psi _{nt_{-}}^{j}$ is given by
equation (\ref{eq:FOC}). This is the same condition as in our baseline case. 
That is ex-ante labor shares should equal expected wage bill shares. 
Note, however, that the stochastic process of $w^j_{nt}$ is different with labor adjustment, so the solution to the ex-ante labor allocation problem will be different than in our baseline case.

@@

\subsubsection{Preferences with Constant Elasticity of Substitution}\label{Sces}

In our baseline model preferences over sectoral goods aggregate in
Cobb-Douglas fashion. In this robustness check we replace equation (\ref%
{aggregate}) by a CES formulation. This requires calibrating $J$ \
\textquotedblleft share\textquotedblright\ parameters, as well as an
elasticity of substitution. Our strategy is to calibrate the share
parameters by matching the share of each final good sector in global expenditure for each year. 
We then look at how our results vary with different
values of the elasticity of substitution.

The results are presented in Table 6. The overall effect of trade on
volatility is quite similar across different specifications of preferences,
though when substitutability is quite low (elasticity of substitution equal
to 0.5) the diversification effect and the sectoral specialization effect
change in offsetting directions

@@

\bibitem{} Nakamura, Emi, and Jón Steinsson. 2008. “Five Facts about Prices: A Reevaluation of Menu Cost Models.” The Quarterly Journal of Economics 123 (4): 1415–64.

\bibitem{} Nakamura, Emi, and Jón Steinsson. 2012. “Lost in Transit: Product Replacement Bias and Pricing to Market.” The American Economic Review 102 (7): 3277–3316.

@@

The goal of this loop is to find the sectoral resource allocations $%
L_{nt}^{j}$ that satisfy %
\begin{equation*}
\frac{L_{nt}^{j}}{L_{nt}}=E_{t-1}\left( \frac{w_{nt}^{j}L_{nt}^{j}}{%
w_{nt}L_{nt}}\right)
\end{equation*}%
where $w_{nt}$ is the average wage. When searching for the
equilibrium value of $L_{nt}^{j}$ the state of the economy is made
up of the deterministic component of the augmented-productivity processes, $%
\bar{Z}_{nt}^{j}$, as well as the previous-period values of the
log-deviation processes representing country, sector, and idyosincratic
shocks, $\lambda _{t-1}^{j}$, $\mu _{nt-1}$, and $%
\epsilon _{nt-1}^{j}$. This state is known both to us and to the
decision maker in the model, as are the autoregressive parameters driving the
shock processes and their variances. Hence, we can draw values from the
distribution of $\lambda _{t}^{j}$, $\mu _{nt}$, and $%
\epsilon _{nt}^{j}$ and combine them with $\bar{Z}_{nt}^{j}$
to create corresponding draws for $Z_{nt}^{j}$. For each iteration
over possible candidates for the $L_{nt}^{j}$s, we thus draw 100
random realizations of the $Z_{nt}^{j}$s, and for each of them we
compute $w_{nt}^{j}$ and $w_{nt}$, and hence $\frac{%
w_{nt}^{j}L_{nt}^{j}}{w_{nt}L_{nt}}$ from the middle loop. Then the
expectation of these wage shares is simply the average across all the draws
of $Z_{nt}^{j}$. The iteration ends when the left hand side and
right hand side are close enough.

@@
